Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-03 examples.tex
+

  \caption{Specification and proof of distilled code for mapping a new page.}

@@

In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not
mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
phyiscal page to back the desired virtual page first requires ensuring
the existence of a memory location for an appropriate L1 table entry.
The code uses a helper function \lstinline|pte\_walkpgdir| (discussed again in Section \ref{sec:traversing}).
\lstinline|pte\_walkpgdir|'s postcondition contains physical pointsto assertions for a partial page table walk reaching the
L1 entry (l1e), ensuring higher levels of the page table exist. 
The postcondition also includes a virtual \emph{PTE}
pointsto ($\mapsto_{\textsf{vpte}}$) allows access to the memory of the L1 entry
via virtual address \lstinline|pte_pointsto|.
A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion 
rather than existentially quantified:
 \[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{vpte}} \;\{\textsf{q}\} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
    \exists \delta\ldotp
	(\lambda \mathit{cr3val}\ldotp
	\sumapaces{\mathit{cr3val}}\delta) \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
\end{array}
\]
This supports rules for accessing memory
at that virtual address, but exposing the physical location being modified
makes this convenient to use for page table modifications, since we must ensure
the modified data is at the specific physical location that will be used as the L1 entry.
This is also guaranteed ($ \ulcorner
\texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$).
% After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
% by the physical memory for the L1 entry that will be used to translate the virtual addresses
% we are mapping, we save it to \textsf{r14} to be updated later in Line 9.

>>>>

  \caption{Specification and proof of code for mapping a new page.}

@@

In Figure \ref{fig:mapping_code}, we see an address ($\vaddr$) currently not
mapped to a page ($\theta \; !!\; \vaddr = \texttt{None}$). Mapping a fresh
phyiscal page to back the desired virtual page first requires ensuring
the existence of a memory location for an appropriate L1 table entry.
The code uses a helper function \lstinline|pte\_walkpgdir| (discussed again in Section \ref{sec:traversing}).
\textsf{pte\_walkpgdir}'s postcondition contains virtual \emph{PTE} pointsto assertions ($\mapsto_{\textsf{vpte}}$) \todo{colin, these are no more plain physical pointstos}  both for ensuring partial page table walk reaching the
L1 entry (l1e) with the justification of higher levels of the page table exist, and for allowing access to the memory of the L1 entry via virtual address.
A PTE points-to is defined just like the normal virtual points-to of Figure \ref{fig:virtualpointstosharing}, except the physical address (here, \textsf{pa}) is explicit in the assertion 
rather than existentially quantified:
 \[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{vpte,q}} \; \paddr \; \vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
    \exists \delta\ldotp
	(\lambda \mathit{cr3val}\ldotp
	\sumapaces{\mathit{cr3val}}\delta) \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
\end{array}
\]
This supports rules for accessing memory
at that virtual address, but exposing the physical location being modified
makes this convenient to use for page table modifications, since we must ensure
the modified data is at the specific physical location that will be used as the L1 entry.
This is also guaranteed ($ \ulcorner
\texttt{addr\_L1}(\vaddr,\entryo) = \paddr\urcorner$).
% After obtaining a virtual address \textsf{pte\_addr} in \textsf{rax} backed 
% by the physical memory for the L1 entry that will be used to translate the virtual addresses
% we are mapping, we save it to \textsf{r14} to be updated later in Line 9.
